### 2020-12-16 CONTRIBUTING.Rmd
Update logo and website

---
title: Contributing to the development of the ropenbals package
author: |
  | **Package website**: **<https://prdm0.github.io/ropenblas>**
  |
  | **Summary**
output: 
  bookdown::github_document2:
    number_sections: true
    toc: true
    toc_depth: 2
toccolor: 'blue'
---

# General information

The **ropenblas** package introduces a simple way, through its functions, to link the library [**OpenBLAS**](https://www.openblas.net/) with the language [**R**](https: //www.r-project.org/). In addition, with ropenblas it is possible to link different versions of R, making it possible to easily switch between the different versions of R compiled.

The package currently works only on GNU/Linux systems and is independent of the official repositories of the user's GNU/Linux distribution. This allows the R user on GNU/Linux systems who is the administrator of the operating system - OS to link the latest versions of R and OpenBLAS as well as switch between versions if they wish.

# Main contributions

The main contributions to the **ropenblas** package are listed below:

  1. Revision of the code to detect possible inconsistent points that may cause the functions provided by the package to fail in some GNU / Linux distribution;
  2. Test the package on different GNU / Linux distributions, reporting possible inconsistencies at <https://github.com/prdm0/ropenblas/issues>;
  3. Improved English for package messages and documentation;
  4. Improved writing of job documentation;
  5. Allow Windows system users to take advantage of the `ropenblas()` function. One suggestion would be to link precompiled binaries from the OpenBLAS library and distribute them in a `inst/` directory.